---
title: "mock egg isolations"
author: "Chenxin Li"
date: "11 - 21 - 2017"
output:
  html_notebook:
    number_sections: yes
    toc: yes
    toc_float: yes
  html_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(emmeans)
library(tidyr)
library(dplyr)
library(readr)
library(readxl)
library(RColorBrewer)
library(svglite)
```

# NENO qubit after bead clean up
```{r}
NENO_Qubit <- read_csv("C:/Users/cxli9/Desktop/Li/PBI/PBI299 (Sundar Lab)/smRNA project/CLVS_NENO/NENO_Qubit.csv")
NENO_Qubit$library <- as.factor(NENO_Qubit$library)
NENO_Qubit$sample_type <- as.factor(NENO_Qubit$sample_type)
```
```{r}
model_neno_q <- lm(qubit_nM ~ sample_type, data = NENO_Qubit)

NENO.test <- t.test(qubit_nM ~ sample_type, NENO_Qubit)
#str(NENO.test)
NENO.test

est_neno_q <- lsmeans(model_neno_q, trt.vs.ctrl~sample_type)
est_neno_q_bar <- as.data.frame(summary(est_neno_q)$lsmeans)

sdn <- NENO_Qubit[1:2,4] %>% var() %>% sqrt()
sen <- sdn/sqrt(2)
sde <- NENO_Qubit[3:8,4] %>% var() %>% sqrt()
see <- sde/sqrt(6)

se_by_group_q <- c(see, sen)
est_neno_q_bar$se_by_group <- se_by_group_q
```

```{r}
ggplot(NENO_Qubit, aes(x= sample_type, y= qubit_nM)) +
  geom_bar(stat = "summary", fun.y = "mean" , aes(fill = sample_type)) +
  geom_point(aes()) +
  ylab("yield(nM)") +
  xlab("sample type")+
  guides(fill =guide_legend(title="sample type")) +
  theme_bw() +
  theme(legend.position = "none") 

 
```

```{r}
2.05/33.08
```
Mock is ~6% of true egg libraries


#qPCR
##Qubit quantification of qPCR standard
Standard is 49% CG, amplicon from plasmid = 109bp, with P5 (20bp) and P7 (24bp) sequence, total = 109 + 20 + 24 = 153bp.  

original concentration of standard is 65ng/ul by qubit
length is 153 bp. 
```{r}
65/660/153 * 10^6
65/660/153 * 10^6 *10^-9 * 6.02*10^23 * 1 *10^-6
```
Molarity = 644nM. 
387502475738 molecules per 1uL. 

##analysis of qPCR results  
```{r}
qPCR_2 <- read_csv("C:/Users/cxli9/Desktop/Li/PBI/PBI299 (Sundar Lab)/smRNA project/CLVS_NENO/Li_2017_12_05_qPCR/qPCR_2.csv")
qPCR_2$sample_type <- as.factor(qPCR_2$sample_type)
qPCR_2$sample_name <- as.factor(qPCR_2$sample_name)
qPCR_2$tech_rep <- as.factor(qPCR_2$tech_rep)
qPCR2_standard <- filter(qPCR_2, sample_type == "standard")
qPCR2_samples <- filter(qPCR_2, sample_type != "standard")

```


```{r}
standard_model2 <- lm(log10(molecule)~Cq, data = qPCR2_standard)
summary(standard_model2)

ggplot(qPCR2_standard, aes(x = Cq,y = log10(molecule))) +
  geom_point(aes(shape = tech_rep),size = 3, alpha = .8) +
  geom_smooth(method='lm') +
  theme_bw() +
  labs(x = "Number of Cycles", y = "log10 Molecule Count") +
  guides(shape =guide_legend(title="technical reps"))
```
log10(molecule count) = -0.272658 * Cq + 10.845645, R^2 = 0.9976.
```{r}
coefs <- summary(standard_model2)
#str(coefs)
coefs$coefficients
intercept <- coefs$coefficients[1,1] %>% signif(digits = 3)
slope <- coefs$coefficients[2,1] %>% signif(digits = 3)
```

```{r}
cor <- cor.test(qPCR2_standard$Cq, log10(qPCR2_standard$molecule))

ggplot(qPCR2_standard, aes(x= Cq, y= log10(molecule))) +
  geom_vline(xintercept = -Inf, size = 1.5) +
  geom_hline(yintercept = -Inf, size = 1.5) +
  geom_smooth(method='lm') +
  geom_point(aes(color = tech_rep),size = 3, alpha = .8)+
  labs(color = "technical reps",
       y = "log10(molecules)") +
  annotate(geom = "text", 
           label = paste("cor = ", cor$estimate %>% signif(digits = 3), sep = ""), 
           x = 20, y = 8.5, size = 5, fontface = "bold") +
  annotate(geom = "text", 
           label = paste( "R^2 = ", 
                         coefs$adj.r.squared %>% signif(digits = 3), sep = ""), 
           x = 20, y = 8, size = 5, fontface = "bold") +
  ggtitle( paste("log10(molecules) = ", 
                         slope, "* Cq + ", intercept, sep = "")) +
  
  theme_minimal() +
  theme(text = element_text(size = 14, face="bold")) +
  theme(axis.text.y = element_text(colour="black")) +
  theme(axis.text.x = element_text(colour="black")) +
  theme(legend.position = c(0.2, 0.2))

ggsave(filename = "qPCR_standard_curve.svg", width = 4.5, height = 4.5)
ggsave(filename = "qPCR_standard_curve.png", width = 4.5, height = 4.5)
```


```{r}
# inspect variation among technical reps

ggplot(qPCR2_samples, aes(x = sample_type, y = Cq)) + 
  geom_point(aes(color = sample_name), position = position_dodge(0.3), alpha = 0.8, size = 2)+
  guides(color =guide_legend(title="library"))+
  xlab("sample type") +
  ylab("Cq")+
  theme_bw() +
  theme(text = element_text(size = 14, face="bold")) +
  theme(axis.text.y = element_text(colour="black")) +
  theme(axis.text.x = element_text(colour="black")) 
```

```{r}
qPCR2_samples <- qPCR2_samples %>% mutate(log.counts = predict(standard_model2, newdata= qPCR2_samples)) %>% mutate(molecule = 10^log.counts)

qPCR2_samples_grouped <-  aggregate(molecule ~ sample_type + sample_name, qPCR2_samples, FUN = mean)

t.test(qPCR2_samples_grouped$molecule ~ qPCR2_samples_grouped$sample_type)
mean_neno <- mean(filter(qPCR2_samples_grouped, sample_type == "no egg no ovary")$molecule)
mean_egg <- mean(filter(qPCR2_samples_grouped, sample_type =="egg")$molecule)
mean_neno/(6.02*10^23)*10^6*10^9*10
mean_egg/(6.02*10^23)*10^6*10^9*10

se_neno <- sd(filter(qPCR2_samples_grouped, sample_type == "no egg no ovary")$molecule)/sqrt(2)
se_egg <- sd(filter(qPCR2_samples_grouped, sample_type == "egg")$molecule)/sqrt(6)

sample_type <- c("egg", "no egg no ovary")
means <- c(mean_egg, mean_neno)
SEM <- c(se_egg, se_neno)
qPCR_mean <- data.frame(sample_type, means, SEM)
```

```{r}
#qPCR2_samples_grouped
ggplot(qPCR2_samples_grouped, aes(x = sample_type, y = molecule)) +
  geom_vline(xintercept = -Inf, size = 1.5) +
  geom_hline(yintercept = -Inf, size = 1.5) +
  geom_bar(stat = "summary", fun.y = "mean" , aes(fill = sample_type), alpha = 0.8) +
  stat_summary(geom = "errorbar",  fun.data = "mean_cl_boot", aes(), width = 0.3) +
  #geom_point(aes(), color = "grey40", position = position_jitter(0.1)) +
  ylab("number of molecules") +
  xlab("")+
  scale_x_discrete(label = c("egg\nn = 6 biological reps", "mock\nn = 2 biological reps")) +
  theme_minimal() +
  theme(legend.position = "none") +
  theme(text = element_text(size = 14, face="bold")) +
  theme(axis.text.y = element_text(colour="black")) +
  theme(axis.text.x = element_text(colour="black")) 

ggsave(filename = "qPCR_quantification.svg", width = 4.5, height = 4.5)
ggsave(filename = "qPCR_quantification.png", width = 4.5, height = 4.5)
```

```{r}
mean_neno
mean_egg
mean_neno/mean_egg
```
4.2*10^7 / 8.5*10^8 = 0.0496.
5% of the reads could be from ovary. 


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
